```{=typst}
#pagebreak(weak: true)
```

# Analysis configuration & data sources

```{r settings_pdf_text}
#| echo: false
#| eval: true

gencode_meta <- cps_report$settings$reference_data$source_metadata |>
  dplyr::filter(source_abbreviation == "gencode")
gencode_label <- paste0(
  "GENCODE ",
  if (isTRUE(as.logical(cps_report$settings$conf$vep$vep_gencode_basic)))
    "basic set" else "all transcripts",
  " (v", gencode_meta$source_version, ")"
)
```

This report was generated with _CPSR_ v`r cps_report$settings$software$cpsr_version`
(backed by _PCGR_ v`r cps_report$settings$software$pcgr_version`).

```{r settings_pdf_table}
#| echo: false
#| eval: true

conf <- cps_report$settings$conf
data.frame(
  Setting = c(
    "Genome assembly",
    "Max gnomAD MAF (non-ClinVar variants)",
    "ClinVar trust level",
    "Secondary findings (ACMG)",
    "Pharmacogenomic findings",
    "Sample sex",
    "Non-protein-coding variants included",
    "VEP transcript set",
    "VEP transcript pick order",
    "VEP regulatory annotation"),
  Value = c(
    cps_report$settings$genome_assembly,
    conf$variant_classification$max_af_gnomad,
    conf$variant_classification$clinvar_trust_level,
    as.logical(conf$variant_classification$secondary_findings),
    as.logical(conf$variant_classification$pgx_findings),
    if (is.null(conf$sample_properties$sex)) "UNKNOWN" else
      conf$sample_properties$sex,
    as.logical(conf$other$show_noncoding),
    gencode_label,
    stringr::str_replace_all(conf$vep$vep_pick_order, ",", ", "),
    as.logical(conf$vep$vep_regulatory))) |>
  dplyr::mutate(Value = as.character(.data$Value)) |>
  tinytable::tt(width = c(0.45, 0.55)) |>
  tinytable::format_tt(escape = TRUE) |>
  cpsr:::style_tt_header_pdf()
```

## Virtual gene panel

```{r virtual_panel_pdf}
#| echo: false
#| output: asis
#| eval: true

## genes of the virtual panel (as shown in the HTML report): variants are
## classified and reported for these genes only
vp_conf <- cps_report$settings$conf$gene_panel
vp_genes <- vp_conf$panel_genes
if ("PRIMARY_TARGET" %in% colnames(vp_genes)) {
  vp_genes <- vp_genes[vp_genes$PRIMARY_TARGET %in% TRUE, , drop = FALSE]
}
vp_genes <- vp_genes |>
  dplyr::distinct(.data$SYMBOL, .keep_all = TRUE) |>
  dplyr::arrange(.data$SYMBOL)
vp_desc <- vp_conf$description
if (!is.null(vp_conf$description_trait) &&
    !any(vp_conf$panel_id %in% c("-1", "0"))) {
  vp_desc <- paste0(vp_desc, ": ", vp_conf$description_trait)
}
## tiles as in the HTML report (teal: genes from a combination of panels)
vp_tiles <- vapply(seq_len(NROW(vp_genes)), function(i) {
  fill <- if (isTRUE(vp_genes$CONFIDENCE_LEVEL[i] == 5)) "#007a74" else "#2c313c"
  paste0('box(width: 100%, fill: rgb("', fill, '"), radius: 2pt, ',
         'inset: (y: 3.5pt), align(center, text(fill: white, size: 7pt, ',
         'weight: "bold")[', vp_genes$SYMBOL[i], ']))')
}, character(1))

cat(
  "The *virtual gene panel* defines the genes for which variants are ",
  "classified and reported (it does not reflect which genes were ",
  "sequenced). Panel: **", cpsr:::doc_note_pdf(vp_desc), "** - **",
  NROW(vp_genes), "** genes",
  if (isTRUE(as.logical(vp_conf$diagnostic_grade_only)))
    " (diagnostic-grade genes only)",
  ".\n\n",
  "```{=typst}\n#block(width: 100%, inset: 8pt, radius: 3pt, ",
  "stroke: 0.6pt + luma(200), fill: luma(250))[\n",
  "#grid(columns: 10, gutter: 3pt,\n",
  paste(vp_tiles, collapse = ",\n"),
  ")\n]\n```\n\n",
  sep = "")
```

## Data sources and versions

```{r data_versions_pdf}
#| echo: false
#| output: asis
#| eval: true

## datasets as listed in the HTML report (Documentation - Dataset
## versions), in two columns with a smaller font
ref_datasets <- cps_report$settings$reference_data$source_metadata
if ("wflow" %in% colnames(ref_datasets)) {
  ref_datasets <- ref_datasets[
    stringr::str_detect(ref_datasets$wflow, "cpsr"), , drop = FALSE]
}
ref_datasets <- ref_datasets[
  ref_datasets$source_abbreviation != "oncokb" &
    !(ref_datasets$source_abbreviation == "dbnsfp" &
        ref_datasets$source_type %in% "gene"), , drop = FALSE]

## markdown special characters in free text
md_escape <- function(x) {
  stringr::str_replace_all(x, "([\\\\`*_\\[\\]<>#@$])", "\\\\\\1")
}
entries <- vapply(seq_len(NROW(ref_datasets)), function(i) {
  d <- ref_datasets[i, ]
  url <- d$source_url
  ## outdated ACMG URL (as in the HTML report)
  if (d$source_abbreviation == "acmg_sf") {
    url <- "https://pubmed.ncbi.nlm.nih.gov/37347242/"
  }
  name <- if (!is.na(url)) paste0("[", md_escape(d$source), "](", url, ")") else
    md_escape(d$source)
  version <- if (!is.na(d$source_version))
    paste0(" (**", md_escape(d$source_version), "**)") else ""
  license <- if (is.na(d$source_license)) "" else {
    ## licenses that require agreements for commercial use in bold
    lic <- md_escape(d$source_license)
    if (d$source_abbreviation %in% c("cgc", "gepa", "dbnsfp"))
      lic <- paste0("**", lic, "**")
    if (!is.na(d$source_license_url))
      lic <- paste0("[", lic, "](", d$source_license_url, ")")
    paste0(" - ", lic)
  }
  paste0("- **", name, "**", version, " - ",
         md_escape(d$source_description), license)
}, character(1))

## two balanced columns (split by text length)
cum_len <- cumsum(nchar(entries))
n_left <- max(1, which.min(abs(cum_len - cum_len[length(cum_len)] / 2)))
col_left <- entries[seq_len(n_left)]
col_right <- entries[-seq_len(n_left)]

cat(
  "PCGR data bundle version: **",
  cps_report$settings$reference_data$version, "**. ",
  "Resources with licenses in bold require specific licensing agreements ",
  "for commercial (non-research) use.\n\n",
  "```{=typst}\n#block[\n#set text(size: 7pt, hyphenate: false)\n",
  "#set par(leading: 0.45em, justify: false)\n",
  "#set list(spacing: 0.45em)\n",
  "#grid(columns: (1fr, 1fr), column-gutter: 1.4em)[\n```\n\n",
  paste(col_left, collapse = "\n"),
  "\n\n```{=typst}\n][\n```\n\n",
  paste(col_right, collapse = "\n"),
  "\n\n```{=typst}\n]\n]\n```\n\n",
  sep = "")
```
