## Variant classification

This section displays all variant classifications, considering both ClinVar and CPSR as the assertion authorities. Use the filters below to explore variants by gene, clinical significance, consequence, and other attributes. Click any row to expand and view detailed annotations, evidence, and classification rationale.

<br>

::: {.callout-note title="Assertion authority — colour legend" collapse="true"}

Each variant row is marked with a coloured vertical bar on the left side of the
_Alteration_ column, indicating which authority determined the final reported
classification:

<span style='display:inline-block; border-left:6px solid #0277bd; padding-left:8px; color:#0277bd; font-weight:bold;'>ClinVar</span>
&nbsp;— classification is sourced from an existing ClinVar record that meets the
configured *clinvar_trust_level* threshold

<span style='display:inline-block; border-left:6px solid #e65100; padding-left:8px; color:#e65100; font-weight:bold;'>CPSR</span>
&nbsp;— classification was assigned *de novo* by the CPSR rule-based ACMG/AMP
scoring engine (either no ClinVar record exists, or the existing record does not
meet the trust threshold).

:::

```{r setup_classification_unified}
#| echo: false
#| output: false
#| eval: true

n_cpg_non_sf <- cps_report$content$snv_indel$v_stat_cpg_non_sf$n
max_rows <- 1000
variants_shared <- list()

if (n_cpg_non_sf > 0) {

  # Prepare unified dataset (all variants, all classes, both sources)
  unified_variants <- cpsr::prepare_unified_all_variants(
    cps_report = cps_report,
    max_rows = max_rows)

  # Count summary
  n_clinvar_total <- sum(unified_variants$ASSERTION_AUTHORITY == "ClinVar")
  n_cpsr_total <- sum(unified_variants$ASSERTION_AUTHORITY == "CPSR")

  summary_text <- htmltools::HTML(paste0(
    "<strong>Total variants: ", nrow(unified_variants), "</strong>",
    " &nbsp;&nbsp; ",
    "<span style='display:inline-block; border-left:6px solid #0277bd;",
    " padding-left:6px; color:#0277bd; font-weight:bold;'>ClinVar-asserted",
    ": ", n_clinvar_total,
    " </span>&nbsp;&nbsp; ",
    "<span style='display:inline-block; border-left:6px solid #e65100;",
    " padding-left:6px; color:#e65100; font-weight:bold;'>CPSR-asserted",
    ": ", n_cpsr_total," </span>"
  ))

  # Create shared crosstalk data
  variants_shared <- list()
  variants_shared[['coding']] <- crosstalk::SharedData$new(
    unified_variants |> dplyr::filter(.data$CODING_STATUS == "coding"),
    key = ~VARKEY_CLASSIFICATION
  )

  variants_shared[['noncoding']] <- crosstalk::SharedData$new(
    unified_variants |> dplyr::filter(.data$CODING_STATUS == "noncoding"),
    key = ~VARKEY_CLASSIFICATION
  )

  ## DP_CONTROL is set to -1 when CPSR is run without a matched control
  ## sample (i.e. depth is unknown for all variants in that case) - only
  ## flag variants as low-depth when a real (non-sentinel) depth value is
  ## present for at least one variant
  dp_threshold <- 10
  has_control_depth <- "DP_CONTROL" %in% colnames(unified_variants) &&
    any(unified_variants$DP_CONTROL >= 0, na.rm = TRUE)
}

```

```{r low_depth_legend}
#| echo: false
#| output: asis
#| eval: !expr n_cpg_non_sf > 0

if (has_control_depth) {
  cat(paste0(
    "\n::: {.callout-note title='Low sequencing depth' collapse='true'}\n\n",
    "Variants with a low control-sample sequencing depth (< ", dp_threshold,
    "X) are flagged with an orange marker on the genotype - interpret their ",
    "classification with caution.\n\n:::\n"
  ))
}
```

```{r no_variants_classification}
#| echo: false
#| eval: !expr cps_report$content$snv_indel$v_stat_cpg_non_sf$n == 0

htmltools::HTML(paste0(
  "<p style='color:#888; font-style:italic;'>",
  "No variants were identified in cancer predisposition genes ",
  "(virtual panel genes, excluding secondary findings).</p>"
))
```


::: {.panel-tabset}

### Coding variants

```{r filters_coding_unified}
#| echo: false
#| output: asis
#| eval: !expr cps_report$content$snv_indel$v_stat_cpg_non_sf$n_coding > 0

cpsr::create_unified_variant_filters(
  shared_data = variants_shared$coding,
  cps_report = cps_report,
  coding_status = "coding"
)


```

```{r table_unified_coding}
#| echo: false
#| eval: !expr cps_report$content$snv_indel$v_stat_cpg_non_sf$n_coding > 0

cpsr::create_unified_variant_reactable(
  shared_data = variants_shared$coding,
  primary_cols =
    c("SYMBOL",
      "gnomADg_AF",
      "CONSEQUENCE",
      "ALTERATION",
      "CLASSIFICATION",
      "GENOTYPE",
      "VARKEY_CLASSIFICATION",
      "ASSERTION_AUTHORITY"),
  color_palette = cpsr::color_palette,
  flag_low_depth = has_control_depth,
  dp_threshold = dp_threshold
)
```

### Non-coding variants

```{r filters_noncoding_unified}
#| echo: false
#| output: asis
#| eval: !expr cps_report$content$snv_indel$v_stat_cpg_non_sf$n_noncoding > 0

cpsr::create_unified_variant_filters(
  shared_data = variants_shared$noncoding,
  cps_report = cps_report,
  coding_status = "noncoding"
)
```

```{r table_unified_noncoding}
#| echo: false
#| output: asis
#| eval: !expr cps_report$content$snv_indel$v_stat_cpg_non_sf$n_noncoding > 0

cpsr::create_unified_variant_reactable(
  shared_data = variants_shared$noncoding,
  primary_cols =
    c("SYMBOL",
      "gnomADg_AF",
      "CONSEQUENCE",
      "ALTERATION",
      "CLASSIFICATION",
      "GENOTYPE",
      "VARKEY_CLASSIFICATION",
      "ASSERTION_AUTHORITY"),
  color_palette = cpsr::color_palette,
  flag_low_depth = has_control_depth,
  dp_threshold = dp_threshold
)

```

:::
