## Documentation

This report is intended for interpretation of inherited DNA variants implicated with cancer susceptibility and inherited cancer syndromes. The set of classified variants are limited to those overlapping a user-defined set of known cancer predisposition target genes (i.e. _virtual panel_), for which the report lists __ONLY__ those variants that are

1. Previously classified in ClinVar (five-level clinical significance scheme: pathogenic/likely pathogenic/VUS/likely benign/benign), or
2. Coding, non-ClinVar-classified variants with germline population frequency below the user-defined threshold, i.e. 
    * Minor allele frequency (MAF) < `r cps_report[['settings']][['conf']][['variant_classification']][['maf_upper_threshold']]`) in the user-defined population set in the gnomAD database
    * The set of novel, _non-ClinVar_ variants are classified using CPSR's dedicated variant classification algorithm, applying the same five-level clinical significance scheme as found in ClinVar

<br>

::: {.panel-tabset}

### Gene/variant annotation resources

The variant interpretation procedures performed in this cancer genome report are relying upon multiple tools and knowledge resources, as outlined below. 

::: {.callout-important title="Licensing note"}

Some resources, specifically highlighted in _Dataset versions_ below, demand specific licensing agreements if you plan to utilize their data (and thus this report) in a commercial, non-research setting.

:::

::: {.callout-note title="Dataset versions" collapse="true"}

* __PCGR databundle version__

  * __`r cps_report$settings$reference_data$version`__

* __Databases/datasets__ 
```{r list_db}
#| echo: false
#| output: asis

i <- 1
ref_datasets <- cps_report$settings$reference_data$source_metadata
for(i in 1:NROW(ref_datasets)){
  source_full <- ref_datasets[i,"source"]
  source <- ref_datasets[i,"source_abbreviation"]
  source_type <- ref_datasets[i, "source_type"]
  description <- ref_datasets[i,"source_description"]
  url <- ref_datasets[i,"source_url"]
  version <- ref_datasets[i,"source_version"]
  license <- ref_datasets[i, "source_license"]
  license_url <- ref_datasets[i, "source_license_url"]
  
  ## temporary fix for outdated ACMG URL (geneOncoX needs update):
  if(source == "acmg_sf"){
    url <- "https://pubmed.ncbi.nlm.nih.gov/37347242/"
  }
  wflow <- ref_datasets[i, "wflow"]
  if(!(stringr::str_detect(
    wflow,"cpsr"))){
    next
  }
  
  if(source == "dbnsfp" & source_type == "gene"){
    next
  }
  
  s <- ""
  if(is.na(version)){
    if(!is.na(license_url)){
      s <- paste0("    * [", source_full, "](", url, ") - ", description, " - [",
                  license,"](", license_url, ")")
    }else{
      s <- paste0("    * [", source_full, "](", url, ") - ", description, " - ",
                  license)
    }
  }else{
    if(!is.na(license_url)){
      if(source == "cgc" | source == "gepa" | source == "dbnsfp"){
        s <- paste0("    * [", source_full, "](", url, ") - ", description, " (<b>", version, "</b>)",
                    " - [<b>", license,"</b>](", license_url, ")")
      }else{
        s <- paste0("    * [", source_full, "](", url, ") - ", description, " (<b>", version, "</b>)",
                    " - [", license,"](", license_url, ")")
      }
    }else{
       s <- paste0("    * [", source_full, "](", url, ") - ", description, " (<b>", version, "</b>)",
                  " - ", license)
    }
  }
  if(s != ""){
    cat(s, sep = "\n")
  }
}

```

:::

### Variant classification


::: {.callout-note title="Synopsis" collapse="false"}

All coding, <b>non-ClinVar</b> variants in the set of cancer predisposition genes are classified according to a <i>five-level pathogenicity scheme</i> (*CPSR_CLASSIFICATION* in the interactive variant classification tables). The scheme has the same five levels as those employed by ClinVar, e.g. pathogenic/likely pathogenic/VUS/likely benign/benign. The classification performed by CPSR is rule-based, implementing a refined set of proposed **ACMG/AMP criteria**, many of which were originally outlined in [SherLoc](https://www.invitae.com/en/variant-classification/) ([Nykamp et al., Genetics in Medicine, 2017](https://www.ncbi.nlm.nih.gov/pubmed/28492532)). Important attributes of cancer predisposition genes, such as mode of inheritance and mechanism of disease (loss-of-function), have been harvested from [Genomics England PanelApp](https://panelapp.genomicsengland.co.uk/), [Maxwell et al., Am J Hum Genet, 2016](https://www.ncbi.nlm.nih.gov/pubmed/27153395), and [Huang et al., Cell, 2018](https://www.ncbi.nlm.nih.gov/pubmed/29625052)

The ACMG/AMP criteria listed in @nte-table-criteria form the basis for the *CPSR_CLASSIFICATION* variable provided by CPSR. Specifically, the <i>score</i> column indicates how much each evidence item contribute to either of the two pathogenicity poles (positive values indicate pathogenic support, negative values indicate benign support). Evidence score along each pole ('B' and 'P') are aggregated, and represented through the *CPSR_PATHOGENICITY_SCORE* variable. See @nte-threshold-calibration on how CPSR establishes optimal thresholds for conversion of pathogenicity scores to categorical variant classification levels. 

:::

::: {.callout-note title="Implemented evidence criteria - ACMG/AMP" collapse="true" #nte-table-criteria}

In the table below, a detailed description of all evidence criteria that are currently used for variant classification in CPSR (green elements indicate criteria that contribute with a benign effect, red elements contribute with a pathogenic effect):<br><br>


```{r acmg_evidence}
#| echo: false
#| eval: true

data <- dplyr::filter(
  cpsr::acmg[["evidence_codes"]],
  cpsr_evidence_code != "ACMG_BS2_1" &
    cpsr_evidence_code != "ACMG_BS2_2" &
    cpsr_evidence_code != "ACMG_BS2_3") |>
  dplyr::select(-category) |>
  dplyr::rename(pole = pathogenicity_pole, 
                category = category_long, 
                score = path_score) |>
  dplyr::arrange(pole)

DT::datatable(data,
  escape = F, extensions = c("Buttons", "Responsive"),
  options = list(
    pageLength = 10,
    buttons = c("csv", "excel"),
    dom = "Bfrtip"
  )
) |>
  DT::formatStyle("cpsr_evidence_code", "pole",
    color = "white",
    backgroundColor = DT::styleEqual(c("P", "B"), 
                                     c("#9E0142", "#077009"))
  )
```


:::

::: {.callout-note title="Calibration of classification thresholds - ACMG/AMP" collapse="true" #nte-threshold-calibration}

How does CPSR assign classification of novel variants (<i>P, LP, VUS, LB, B</i>) from the aggregated variant pathogenicity score (<i>CPSR_PATHOGENICITY_SCORE</i>)?

Optimal thresholds for conversion of pathogenicity scores to categorical variant classification were calibrated through high-quality ClinVar-classified variants (__minimum two gold stars__ with respect to review status) in a core set of cancer predisposition genes (see details in [CPSR manuscript](https://doi.org/10.1002/ijc.33749), and plot below). 

```{r classification_plot}
#| echo: false
#| eval: true

classification_calibration_plot_fname_path <- 
  system.file("extdata", "CPSR_classification_score_thresholds.png", 
              package = "cpsr")
```

<img src="`r classification_calibration_plot_fname_path`" title="" alt="" align="center" width="950" height="420"/>

<br>

The following thresholds are currently used to assign classifications based on pathogenicity scores:

<br>

`r knitr::kable(cpsr::acmg[['score2tier']])`

<br>

:::

:::

