## Documentation

`r cpsr::cpsr_intro_doc_note()`

<br>

::: {.panel-tabset}

### Gene/variant annotation resources

The variant interpretation procedures performed in this cancer genome report are relying upon multiple tools and knowledge resources, as outlined below. 

::: {.callout-important title="Licensing note"}

Some resources, specifically highlighted in _Dataset versions_ below, demand specific licensing agreements if you plan to utilize their data (and thus this report) in a commercial, non-research setting.

:::

::: {.callout-note title="Dataset versions" collapse="true"}

* __PCGR databundle version__

  * __`r cps_report$settings$reference_data$version`__

* __Databases/datasets__ 
```{r list_db}
#| echo: false
#| output: asis

i <- 1
ref_datasets <- cps_report$settings$reference_data$source_metadata
for(i in 1:NROW(ref_datasets)){
  source_full <- ref_datasets[i,"source"]
  source <- ref_datasets[i,"source_abbreviation"]
  source_type <- ref_datasets[i, "source_type"]
  description <- ref_datasets[i,"source_description"]
  url <- ref_datasets[i,"source_url"]
  version <- ref_datasets[i,"source_version"]
  license <- ref_datasets[i, "source_license"]
  license_url <- ref_datasets[i, "source_license_url"]
  
  ## temporary fix for outdated ACMG URL (geneOncoX needs update):
  if(source == "acmg_sf"){
    url <- "https://pubmed.ncbi.nlm.nih.gov/37347242/"
  }
  wflow <- ref_datasets[i, "wflow"]
  if(!(stringr::str_detect(
    wflow,"cpsr"))){
    next
  }
  
  if(source == "oncokb"){
    next
  }
  
  if(source == "dbnsfp" & source_type == "gene"){
    next
  }
  
  s <- ""
  if(is.na(version)){
    if(!is.na(license_url)){
      s <- paste0("    * [", source_full, "](", url, ") - ", description, " - [",
                  license,"](", license_url, ")")
    }else{
      s <- paste0("    * [", source_full, "](", url, ") - ", description, " - ",
                  license)
    }
  }else{
    if(!is.na(license_url)){
      if(source == "cgc" | source == "gepa" | source == "dbnsfp"){
        s <- paste0("    * [", source_full, "](", url, ") - ", description, " (<b>", version, "</b>)",
                    " - [<b>", license,"</b>](", license_url, ")")
      }else{
        s <- paste0("    * [", source_full, "](", url, ") - ", description, " (<b>", version, "</b>)",
                    " - [", license,"](", license_url, ")")
      }
    }else{
       s <- paste0("    * [", source_full, "](", url, ") - ", description, " (<b>", version, "</b>)",
                  " - ", license)
    }
  }
  if(s != ""){
    cat(s, sep = "\n")
  }
}

```

:::

### Variant classification

::: {.callout-note title="Synopsis" collapse="true"}

`r cpsr::variant_classification_synopsis_doc_note(quarto = TRUE)`

:::

::: {.callout-note title="Variant loss-of-function annotation" collapse="true" #nte-lof-annotation}

`r pcgrr::lof_doc_note()`

:::

::: {.callout-note title="Implemented evidence criteria - ACMG/AMP" collapse="true" #nte-table-criteria}

In the table below, a detailed description of all evidence criteria that are currently used for variant classification in CPSR (green elements indicate criteria that contribute with a benign effect, red elements contribute with a pathogenic effect):<br><br>


```{r acmg_evidence}
#| echo: false
#| eval: true

data <- 
  cpsr::acmg[["evidence_codes"]] |>
    #dplyr::filter(
    #cpsr_evidence_code != "ACMG_PP2" &
    #cpsr_evidence_code != "ACMG_PM4") |>
  dplyr::select(-category) |>
  dplyr::rename(pole = pathogenicity_pole, 
                category = category_long, 
                score = path_score) |>
  dplyr::arrange(pole)

DT::datatable(data,
  escape = F, 
  extensions = c("Buttons", "Responsive"),
  options = list(
    pageLength = 10,
    buttons = c("csv", "excel"),
    dom = "Bfrtip"
  )) |>
  DT::formatStyle(
    "cpsr_evidence_code", "pole",
    color = "white",
    backgroundColor = DT::styleEqual(
      c("P", "B"), 
      c("#9E0142", "#077009"))
  )
```


:::


::: {.callout-note title="Use of gnomAD in variant classification - ACMG/AMP" collapse="true" #nte-gnomad}

`r cpsr::gnomad_germline_doc_note()`

:::

::: {.callout-note title="Calibration of classification thresholds - ACMG/AMP" collapse="true" #nte-threshold-calibration}

`r cpsr::classification_thresholds_intro_doc_note()`

```{r classification_plot}
#| echo: false
#| eval: true

classification_calibration_plot_fname_path <-
  system.file("extdata", "CPSR_classification_score_thresholds.png",
              package = "cpsr")
```

<img src="`r classification_calibration_plot_fname_path`" title="" alt="" align="center" width="950" height="420"/>

<br>

The following thresholds are currently used to assign classifications based on pathogenicity scores:

<br>

`r knitr::kable(cpsr::acmg[['score2tier']])`

<br>

`r cpsr::classification_thresholds_outro_doc_note()`

<br>

:::

:::

