## Variant Classification

This section displays all variant classifications, considering both ClinVar and CPSR as the assertion authorities. Use the filters below to explore variants by gene, clinical significance, consequence, and other attributes. Click any row to expand and view detailed annotations, evidence, and classification rationale.

<br>

```{r setup_classification_unified}
#| echo: false
#| output: false
#| eval: true

# Configuration
max_rows <- 1000

# Prepare unified dataset (all variants, all classes, both sources)
unified_variants <- cpsr::prepare_unified_all_variants(
  cps_report = cps_report,
  max_rows = max_rows)

# Count summary
n_clinvar_total <- sum(unified_variants$ASSERTION_AUTHORITY == "ClinVar")
n_cpsr_total <- sum(unified_variants$ASSERTION_AUTHORITY == "CPSR")

summary_text <- htmltools::HTML(paste0(
  "<strong>Total variants: ", nrow(unified_variants), "</strong>",
  " &nbsp;&nbsp; ",
  "<span style='display:inline-block; border-left:6px solid #0277bd;",
  " padding-left:6px; color:#0277bd; font-weight:bold;'>ClinVar-asserted",
  ": ", n_clinvar_total,
  " </span>&nbsp;&nbsp; ",
  "<span style='display:inline-block; border-left:6px solid #e65100;",
  " padding-left:6px; color:#e65100; font-weight:bold;'>CPSR-asserted",
  ": ", n_cpsr_total," </span>"
))


# Create shared crosstalk data
variants_shared <- list()
variants_shared[['coding']] <- crosstalk::SharedData$new(
  unified_variants |> dplyr::filter(.data$CODING_STATUS == "coding"),
  key = ~VARKEY_CLASSIFICATION
)

variants_shared[['noncoding']] <- crosstalk::SharedData$new(
  unified_variants |> dplyr::filter(.data$CODING_STATUS == "noncoding"),
  key = ~VARKEY_CLASSIFICATION
)

```

`r summary_text`

::: {.callout-note title="Note - output limitations" collapse="true"}
For performance, a maximum of `r max_rows` variants per source (ClinVar/CPSR) are displayed. Variants are pre-sorted by clinical significance and review status to prioritize high-evidence variants.
:::

<br>

::: {.panel-tabset}

### Coding variants 

```{r filters_coding_unified}
#| echo: false
#| output: asis

cpsr::create_unified_variant_filters(
  shared_data = variants_shared$coding,
  cps_report = cps_report,
  coding_status = "coding"
)


```

```{r table_unified_coding}
#| echo: false

cpsr::create_unified_variant_reactable(
  shared_data = variants_shared$coding,
  primary_cols = 
    c("SYMBOL", 
      "gnomADg_AF",
      "CONSEQUENCE", 
      "ALTERATION", 
      "CLASSIFICATION", 
      "GENOTYPE",
      "VARKEY_CLASSIFICATION", 
      "ASSERTION_AUTHORITY"),
  color_palette = cpsr::color_palette
)
```

### Non-coding variants

```{r filters_noncoding_unified}
#| echo: false
#| output: asis

cpsr::create_unified_variant_filters(
  shared_data = variants_shared$noncoding,
  cps_report = cps_report,
  coding_status = "noncoding"
)
```

```{r table_unified_noncoding}
#| echo: false
#| output: asis
#| eval: true

cpsr::create_unified_variant_reactable(
  shared_data = variants_shared$noncoding,
  primary_cols = 
    c("SYMBOL", 
      "gnomADg_AF",
      "CONSEQUENCE", 
      "ALTERATION", 
      "CLASSIFICATION", 
      "GENOTYPE",
      "VARKEY_CLASSIFICATION", 
      "ASSERTION_AUTHORITY"),
  color_palette = cpsr::color_palette
)

```

:::
