---
format:
  typst:
    include-before-body: typst/_preamble.typ
    toc: true
    toc-depth: 2
    toc-title: "Contents"
    papersize: a4
    margin:
      top: 2.5cm
      bottom: 2.2cm
      left: 2cm
      right: 2cm
    section-numbering: ""
---

```{r global_options}
#| include: false
#| eval: true

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, error = FALSE, eval = TRUE)
options(scipen = 999)

cps_report <- readRDS(file = "<CPSR_REPORT_OBJECT.rds>")
no_content <- isFALSE(as.logical(cps_report$content$snv_indel$eval))
```

```{r cover_page}
#| output: asis
#| echo: false
#| eval: true

# ---------------------------------------------------------------------------
# Resolve KPI values and colors (mirrors logic in cpsr_summary.qmd)
# ---------------------------------------------------------------------------
color_none <- cpsr::color_palette[["none"]]

n_path_clinvar <- 0L
n_path_cpsr    <- 0L
n_vus          <- 0L
n_benign       <- 0L

path_genes_str   <- "None"
sf_genes_str     <- "Not determined"
bm_pgx_genes_str <- "None"

col_path       <- color_none
col_vus        <- color_none
col_benign     <- color_none
col_path_genes <- color_none
col_sf         <- color_none
col_bm_pgx     <- color_none

variants_cpg <- cps_report$content$snv_indel$callset$variant_display$cpg_non_sf

if (NROW(variants_cpg) > 0) {
  n_path_clinvar <- sum(
    variants_cpg$ASSERTION_AUTHORITY == "ClinVar" &
      variants_cpg$CLASSIFICATION %in% c("Pathogenic", "Likely Pathogenic"),
    na.rm = TRUE
  )
  n_path_cpsr <- sum(
    variants_cpg$ASSERTION_AUTHORITY == "CPSR" &
      variants_cpg$CLASSIFICATION %in% c("Pathogenic", "Likely Pathogenic"),
    na.rm = TRUE
  )
  n_vus    <- sum(variants_cpg$CLASSIFICATION == "VUS", na.rm = TRUE)
  n_benign <- sum(
    variants_cpg$CLASSIFICATION %in% c("Benign", "Likely Benign"),
    na.rm = TRUE
  )

  path_genes <- cps_report$content$snv_indel$callset$variant$cpg_non_sf |>
    dplyr::filter(CLASSIFICATION %in% c("Pathogenic", "Likely_Pathogenic")) |>
    dplyr::distinct(SYMBOL) |>
    utils::head(3)

  if (NROW(path_genes) > 0) {
    path_genes_str <- if (NROW(path_genes) == 3) "> 2" else
      paste(path_genes$SYMBOL, collapse = " / ")
    col_path_genes <- cpsr::color_palette$pathogenicity$values[5]
  }
}

if ((n_path_clinvar + n_path_cpsr) > 0)
  col_path <- cpsr::color_palette$pathogenicity$values[5]
if (n_vus > 0)
  col_vus <- cpsr::color_palette$pathogenicity$values[3]
if (n_benign > 0)
  col_benign <- cpsr::color_palette$pathogenicity$values[1]

# Secondary findings
sf_enabled <- isTRUE(
  as.logical(cps_report$settings$conf$variant_classification$secondary_findings)
)
if (sf_enabled) {
  sf_genes_str <- "None"
  sf_genes <- cps_report$content$snv_indel$callset$variant$sf |>
    dplyr::distinct(SYMBOL) |>
    utils::head(3)
  if (NROW(sf_genes) > 0) {
    sf_genes_str <- if (NROW(sf_genes) == 3) "> 2" else
      paste(sf_genes$SYMBOL, collapse = " / ")
    col_sf <- cpsr::color_palette$pathogenicity$values[5]
  }
}

# Biomarkers + PGx
bm_genes  <- cps_report$content$snv_indel$callset$variant$biomarker |>
  dplyr::distinct(SYMBOL) |> utils::head(3)
pgx_genes <- cps_report$content$snv_indel$callset$variant$pgx |>
  dplyr::distinct(SYMBOL) |> utils::head(3)
bm_pgx_genes <- dplyr::bind_rows(bm_genes, pgx_genes) |>
  dplyr::distinct(SYMBOL) |> utils::head(3)
if (NROW(bm_pgx_genes) > 0) {
  bm_pgx_genes_str <- if (NROW(bm_pgx_genes) == 3) "> 2" else
    paste(bm_pgx_genes$SYMBOL, collapse = " / ")
  col_bm_pgx <- cpsr::color_palette[["report"]]
}

# Panel description
gene_panel_desc <- cps_report$settings$conf$gene_panel$description
if (!is.null(cps_report$settings$conf$gene_panel$description_trait) &&
    cps_report$settings$conf$gene_panel$panel_id != "-1" &&
    cps_report$settings$conf$gene_panel$panel_id != "0") {
  gene_panel_desc <- paste0(
    gene_panel_desc, ": ",
    cps_report$settings$conf$gene_panel$description_trait
  )
}

# Emit raw Typst call — values are resolved in R before Typst sees them
cat(glue::glue(
  '```{{=typst}}',
  '#cpsr-cover(',
  '  sample_id:       "{cps_report$settings$sample_id}",',
  '  genome_assembly: "{cps_report$settings$genome_assembly}",',
  '  gene_panel:      "{gene_panel_desc}",',
  '  cpsr_version:    "{cps_report$settings$software$cpsr_version}",',
  '  genes_path:      "{path_genes_str}",',
  '  genes_bm_pgx:    "{bm_pgx_genes_str}",',
  '  genes_sf:        "{sf_genes_str}",',
  '  n_path:          "N = {n_path_clinvar + n_path_cpsr}",',
  '  n_vus:           "N = {n_vus}",',
  '  n_benign:        "N = {n_benign}",',
  '  col_genes_path:  "{col_path_genes}",',
  '  col_bm_pgx:      "{col_bm_pgx}",',
  '  col_sf:          "{col_sf}",',
  '  col_path:        "{col_path}",',
  '  col_vus:         "{col_vus}",',
  '  col_benign:      "{col_benign}",',
  ')',
  '```',
  .sep = "\n"
))
```

```{r classification_cpsr_pdf}
#| child: 'quarto/pdf/cpsr_classification_pdf.qmd'
#| eval: !expr !isTRUE(no_content)
```

```{r secondary_findings_pdf}
#| child: 'quarto/pdf/cpsr_secondary_findings_pdf.qmd'
#| eval: !expr isTRUE(sf_enabled) && !isTRUE(no_content)
```

```{r settings_pdf}
#| child: 'quarto/pdf/cpsr_settings_pdf.qmd'
#| eval: true
```

---

*This report was generated with [CPSR](https://github.com/sigven/cpsr) v`r cps_report$settings$software$cpsr_version` on `r strftime(Sys.time(), format = "%B %d, %Y")`. For research use only.*
